Sensor Clusterization in D-optimal Design in Infinite Dimensional Linear Bayesian Inverse Problems

### 2020-10-06 new_paper.tex
Fails because of ell wtf

\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
One can \textit{think} of $(\obs \fwd \param)_j$ as pointwise
evaluations of a continuous function\footnote{Of course, this is
  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$. Still,
  this approach is in line with previous analyses
  \cite{Karina,AlexanderianGloorGhattas14} that consider sensor
  placement but formulate the inverse problem on Hilbert spaces.}. Such
measurements can be represented using the linear functionals
$\delta_{\x_j} \in C(\Omega)^*$. In a different setting, we may be
able to measure other quantities, e.g.\ some Fourier coefficients as
is the case in MRI applications. Either way, the rows of $\obs$ are
linear functionals and we refer to these as {\it measurement
  vector}s. %% Note that we cannot choose every measurement vector we
%% want, though. For example, we may be restricted only to pointwise
%% evaluations of $\fwd \param$ by the sensors at our disposal, in which
%% case measuring the mean $\ell(u) = \int_{\Omega}u$ (or any non-atomic
%% measure) of a function is not possible.

@@

\begin{restatable}[Characterizing D-optimal designs]{theorem}{main}\label{thm:char}
Denote $\{\lambda_i, \ev_i\}_{i=1}^\infty$ the eigenvalues and
eigenvectors of $\fwd\prcov\fwd^*$, in decreasing eigenvalue order.
Denote $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, chosen
to maximize $\tar(\obs)$. Denote $m$ number of measurements taken and
$\sigma^2$ observation noise variance. Then:
\begin{enumerate}
\item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
\item There exists $k \leq m$ such that $\eta_i = 0$ for $i\geq k+1$.
\item $\sigma^2\lambda_i^{-1} + \eta_i \leq \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
\item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
\end{enumerate}
\end{restatable}

%% We are also used to symmetric optimization problems having
%% symmetric solutions \cite{Waterhouse83}, so symmetric and identical
%% Lagrange multipliers definitely make sense in this context.

\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
\subsection{Overview}
In this section we explain how sensor clusterization may arise. We
first consider a sequential optimal design, where measurements are
chosen sequentially. We then consider the case where all measurements
are chosen at once. Finally, we hypothesize how this may apply to a
real-life inverse problem.

%% illustrated in figure \ref{fig:clusterization} and

\subsection{Sequential Design}
Recall the sequential optimal design scheme from section
\ref{subsec:seq vs sim} and the definition of $\obs_k$ from
\eqref{eq:def obs_k}. The way clusterization can occur is understood
using results of the previous section, specialized to one measurement
each time. By corollary \ref{cor:ame ev}, the first measurement
$\meas_1$ has to be (parallel to) an eigenvector of
$\fwd\prcov\fwd^*$. By theorem \ref{thm:char}, it will be taken in the direction of the
first eigenvector --- $\meas_1 \parallel \ev_1$. As hinted in
section \ref{subsec:seq vs sim}, the posterior becomes the prior for
the next step. Similar reasoning leads to:
\begin{align*}
  \begin{split}
    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
    \meas_2 \parallel \ev_2  &\text{ o.w. }
  \end{split}
\end{align*}
In the first case, we get clusterization. In the second, we consider $\meas_3$:
\begin{align*}
  \begin{split}
    \meas_3 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_3 \text{ (note $\lambda^{-1}_1 < \lambda^{-1}_2$)}  \\
    %\meas_3 \parallel  \ev_2  &\text{ if } \\
    \meas_3 \parallel  \ev_3  &\text{ o.w. }
  \end{split}
\end{align*}
Again, the first case results in clusterization. The sequential design
proceeds in this fashion, where every $\meas_i$ is taken in the
direction of the eigenvector of smallest eigenvalue of the previous
posterior precision operator. Taking enough measurements we will
surely arrive at clusterization, since the prior precision increases
quickly --- in section \ref{subsec:abstract OED} we assumed  $\fwd$
is strongly smoothing, which typpically means eigenvalues of $(\fwd
\prcov\fwd^*)^{-1}$ grow exponentially.

>>>>

\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
One can \textit{think} of $(\obs \fwd \param)_j$ as pointwise
evaluations of a continuous function\footnote{Of course, this is
  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$.}. Such
measurements can be represented using the linear functionals
$\delta_{\x_j} \in C(\Omega)^*$. In a different setting, we may be
able to measure other quantities, e.g.\ some Fourier coefficients as
is the case in MRI applications. Either way, the rows of $\obs$ are
linear functionals and we refer to these as {\it measurement
  vector}s. %% Note that we cannot choose every measurement vector we
%% want, though. For example, we may be restricted only to pointwise
%% evaluations of $\fwd \param$ by the sensors at our disposal, in which
%% case measuring the mean $\ell(u) = \int_{\Omega}u$ (or any non-atomic
%% measure) of a function is not possible.

@@

\begin{restatable}[Characterizing D-optimal designs]{theorem}{char}\label{thm:char}
  Let:
  \begin{enumerate}
  \item $\fwd,\prcov,\obs$ the forward, prior covariance and observation operators, respectively.
  \item $\{\lambda_i\}_{i=1}^\infty$ the eigenvalues of
    $\fwd\prcov\fwd^*$ in decreasing order.
    \item $\{\ev_i\}_{i=1}^\infty$ the corresponding eigenvectors.
  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
    $\obs$ maximizes $\tar(\obs)$.
  \item $m$ the number of measurements taken.
  \item $\sigma^2$ observation noise variance.
  \end{enumerate}
  Then:
  \begin{enumerate}
  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq k+1$.
  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
  \end{enumerate}
\end{restatable}

\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
\subsection{Overview}
In this section we use theorem \ref{thm:char} to explain how sensor
clusterization may arise in various scenarios. We first consider a
sequential optimal design. We then consider the case where all
measurements are chosen at once. Finally, we hypothesize how this may
apply to a real-life inverse problem.

\subsection{Sequential Design}
Recall the sequential optimal design scheme from section
\ref{subsec:seq vs sim} and the definition of $\obs_k$ from
\eqref{eq:def obs_k}. The way clusterization can occur in this
scenario is understood using thorem \ref{thm:char}, specialized to one
measurement each time. In this case $m=1$, so $k=1$, so $\meas_1
\parallel \ev_1$. As hinted in section \ref{subsec:seq vs sim}, the
posterior becomes the prior for the next step. Similar reasoning leads
to:
\begin{align*}
  \begin{split}
    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
    \meas_2 \parallel \ev_2  &\text{ o.w. }
  \end{split}
\end{align*}
In the first case, we get clusterization. In the second, we consider $\meas_3$:
\begin{align*}
  \begin{split}
    %\meas_3 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_3 \text{ (note $\lambda^{-1}_1 < \lambda^{-1}_2$)}  \\
    %\meas_3 \parallel  \ev_2  &\text{ if } \\
    \meas_3 \parallel  \ev_3  &\text{ o.w. }
  \end{split}
\end{align*}
Again, the first case results in clusterization. The sequential design
proceeds in this fashion, where every $\meas_i$ is taken in the
direction of the eigenvector of smallest eigenvalue of the previous
posterior precision operator. Taking enough measurements we will
surely arrive at clusterization, since the prior precision increases
quickly --- in section \ref{subsec:abstract OED} we assumed  $\fwd$
is strongly smoothing, which typpically means eigenvalues of $(\fwd
\prcov\fwd^*)^{-1}$ grow exponentially.
